### 2019-06-18 Report/pca.tex
progress on procrustes write up

\documentclass[12pt]{article}

\begin{document}

\section{Blendshapes and PCA}

\end{document}